---
title: "Week1_data_cleaning"
author: "Emmanuel Okyere Darko"
date: "4/15/2021"
output: statsr:::statswithr_lab
---

## Libraries
```{r, cache=TRUE, warning=FALSE, message=FALSE}
#setwd(paste0(getwd(), "Data/en_US/"))
library("R.utils")
library(NLP)
library(tm)

```

## Getting and Cleaning Data 

Read all txt files in a particular language directory, find the number of lines in each,
the longest line and number of words

2. Question 2: The en_US.twitter.txt has how many lines of text?
```{r data, warning=FALSE, cache=TRUE}
flist <- list.files("Data", pattern = ".*en_.*.txt", recursive = T)

get_info <-lapply(paste("Data", flist, sep = "/"), function(filename){
      ## Check file size 
      fsize = file.size(filename)[1]/1024/1024
      
      ## Read subset of data
      con <- file(filename, "r") #open connection
      all_lines = readLines(con)
      #readLines(con, 5) ## read first 5 lines
      
      ## check number of lines in the file 
      #NROW(readLines(con_en_us_twitter))
      nlines <- sapply(filename, countLines)
      
      ##longest line 
      nchars = lapply(all_lines, nchar)
      longest_line <- which.max(nchars) ## determine index of max
      
      ## split based on one or more white spaces and find length
      nwords <- sum(sapply(strsplit(all_lines, "\\s+"), length)) 
      
      close(con)
      return(c(filename, fsize, nlines, longest_line, nwords))
})

```

### Neatly convert list to dataframe and visualize

```{r df, echo=FALSE}
us_df <- data.frame(matrix(unlist(get_info), nrow=length(get_info), byrow=T))
colnames(us_df) <- c("file", "fsize", "num_of_lines", "longest_line", "nwords")
us_df
```

4. In the en_US twitter data set, if you divide the number of lines where the word "love" (all lowercase) occurs by the number of lines the word "hate" (all lowercase) occurs, about what do you get?

```{r, cache= TRUE, warning=FALSE}
con_en_us_twitter <- file("Data/en_US/en_US.twitter.txt", "r")
## grepL returns logical of each word
love <- grepl(".love.", readLines(con_en_us_twitter), ignore.case = FALSE)

## open new connection after reading
con_en_us_twitter <- file("Data/en_US/en_US.twitter.txt", "r")
hate <- grepl(".hate.", readLines(con_en_us_twitter), ignore.case = FALSE)


close(con_en_us_twitter)
print(sum(love)/ sum(hate))
```

5. The one tweet in the en_US twitter data set that matches the word "biostats" says what?

```{r, cache=TRUE, warning=FALSE}
## open new connection after reading
con_en_us_twitter <- file("Data/en_US/en_US.twitter.txt", "r")
tweet <- readLines(con_en_us_twitter)

## note that grep is a function for finding index
biostats_idx <- grep("biostats", tweet, ignore.case = FALSE)
close(con_en_us_twitter)

tweet[biostats_idx]

```

6. Question 6
How many tweets have the exact characters "A computer once beat me at chess, but it was no match for me at kickboxing". (I.e. the line matches those characters exactly.)

```{r, cache=TRUE, warning=FALSE}
## open new connection after reading
con_en_us_twitter <- file("Data/en_US/en_US.twitter.txt", "r")
tweet <- readLines(con_en_us_twitter)

## note that grep is a function for finding index
idx <- grep("A computer once beat me at chess, but it was no match for me at kickboxing",
            tweet, ignore.case = FALSE)
close(con_en_us_twitter)

tweet[idx]
 
```