---
title: "shiny"
author: "Emmanuel Okyere Darko"
date: "4/10/2021"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
library(lubridate)
```

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

## Mtcars
```{r mtcars, echo=FALSE}

# Define UI for application that draws a histogram
shinyApp(
  ui = fluidPage(

    # Application title
    titlePanel("Mtcars Prediction with splines"),

    # Sidebar with a slider input for number of bins
    sidebarLayout(
        sidebarPanel(
            h3('Slide me'),
            sliderInput("mpgSlider",
                        "Choose MPG to predict:",
                        min = 10,
                        max = 40,
                        value = 20),
            
            #numericInput('id1', 'numbers only', 5, 0, 50, step = 5),
            
            #dateInput('date', 'Enter Date'),
            
            checkboxInput('show_mod1', 'Show Model1', value = T),
            checkboxInput('show_mod2', 'Show Model2', value = T),
            checkboxInput('show_col', 'show groups', value = T),
            
            #submitButton('submit')
            
        ),

        # Show a plot of the generated distribution
        mainPanel(
            
            tabsetPanel(type = "tabs",
                        tabPanel(title = "Tab1", br(), plotOutput("Plot"),
                                 ),
                        
                        tabPanel(title = "Prediction", br(), 
                                 h4("Model 1 Prediction"), textOutput("pred2"),
                                 
                                 h4("Model 2 prediction"), verbatimTextOutput('pred1')
                        )
                        )
          
        )
    )
),

# Define server logic required to draw a histogram
server = (function(input, output) {
    
    mtcars$mpg_spline <- ifelse(mtcars$mpg - 20> 0, mtcars$mpg -20, 0)
    
    fit.lm1 <- lm(hp~mpg, data = mtcars)
    fit.lm2  <- lm(hp~mpg + mpg_spline, data = mtcars)
    
    lm1.pred <- reactive({
        mpgInp <- input$mpgSlider
        predict(fit.lm1, newdata = data.frame(mpg = mpgInp))
    })
    
    lm2.pred <- reactive({
        mpgInp <- input$mpgSlider
        predict(fit.lm2, newdata = data.frame(mpg = mpgInp, 
                                              mpg_spline = ifelse(mpgInp -20 >0, mpgInp -20, 0) 
                                              )
                )
    })

    output$Plot <- renderPlot({ 
        mpgInp <- input$mpgSlider
        with(mtcars, plot(mpg, hp, xlab = "MPG", ylab = "Horsepower",
                          ylim = c(50,340), xlim = c(10,35) 
                          )
             )
        
        if(input$show_mod1){
            abline(fit.lm1, col = "red", lwd = 2)
        }
        if(input$show_col){
            with(mtcars, plot(mpg, hp, xlab = "MPG", ylab = "Horsepower", pch = 17,
                              ylim = c(50,340), xlim = c(10,35), col = as.factor(mtcars$cyl)
                              )
    
            )
        }
        if(input$show_mod2){
            model2_splines <- predict(fit.lm2, 
                                      newdata = data.frame(mpg = 10:35, 
                                            mpg_spline = ifelse(10:35 -20 >0, 10:35 -20, 0)
                                            )
                                      )
            
        lines(10:35, model2_splines, col = 'blue', lwd = 2)
        
        }
        
        points(mpgInp, lm2.pred(), col = 'blue', pch = 16)
        points(mpgInp, lm1.pred(), col = 'red', pch = 16)
    })

    output$pred1 <- renderText({
        lm1.pred()
        })
    
    output$pred2 <- renderText({lm2.pred()})
    

})
)

```

## 
```{r eruptions, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.



