---
title: "Regression_Mtcars"
author: "Emmanuel Okyere Darko"
date: "3/28/2021"
output: statsr:::statswithr_lab
---

## Excutive Summary

This project will focus on a data collection of cars, exploring the relationship
between a set of variables and miles per gallon (MPG) as outcome variable.
Interested questions for this project are 

* Is automatic or manual transmission better for MPG?
* Quantify the MPG difference between manual and automatic transmission

1. Data Description
A data frame with 32 observations on 11 (numeric) variables.

* [, 1]	mpg	Miles/(US) gallon
* [, 2]	cyl	Number of cylinders
* [, 3]	disp	Displacement (cu.in.)
* [, 4]	hp	Gross horsepower
* [, 5]	drat	Rear axle ratio
* [, 6]	wt	Weight (1000 lbs)
* [, 7]	qsec	1/4 mile time
* [, 8]	vs	Engine (0 = V-shaped, 1 = straight)
* [, 9]	am	Transmission (0 = automatic, 1 = manual)
* [,10]	gear	Number of forward gears
* [,11]	carb	Number of carburetors

```{r libraries, warning=FALSE, echo=FALSE}
library(ggplot2)
library(datasets)
data("mtcars")
```

## Exploratory Analysis

```{r cars}
str(mtcars)
```

1. Change data types
* Cylinder is a count variable, therefore convert to factor `4, 6 or 8`
* gear also as factor varible (3,4 or 5)
* Transmission (am) as factor - `(0 = automatic, 1 = manual)`

```{r }
mtcars[,c(2,9,10)] <- sapply(mtcars[,c(2,9,10)] , factor)

```

Since there are no complete overlaps in the boxplot below, Transmission type
appears to have better relationship with mpg

```{r }
with(mtcars, boxplot(mpg~am))
```

## Model

1. We will fit a linear linear model with only cylinder as the predictor varible  
Transmission (am) as factor - `(0 = automatic, 1 = manual)`  

```{r }
fit.cyl <- lm(mpg~am, data= mtcars)
summary(fit.cyl)
```

Transimission type is a significant predictor of MPG with based on the p-values. Automatic transmission 
appears to consume fewer MPG with Manual transmission expected consume and 7.2 more gallons per 1 mile.
However, the low R-Squared shows that approximately `36%` of the variation in the model is explained by 
the Transmission variable and our co-efficients may not be reliable.


2. Let's compare this to a full model

```{r }
fit.full <- lm(mpg~., data= mtcars)
summary(fit.full)
```

The full model gives a relatively high Adjusted R-squared `88.5%` but has less significant variables,
implying that there could be overfitting.


3. Selecting the best model
All numerical variables have at least moderate correlation with mpg
```{r}
library(GGally)
df_cor <- na.omit(mtcars[, c(1,3:8,11)])
ggcorr(df_cor, label = T)
```

We can use the step fubction to select the best model, we will use backward 
stepwise regression, removing one variable a time untill we reach the best model (when AIC stops
decreasing)

```{r}
# To see output, specify trace = True
fit.best <- step(fit.full, direction = "backward", trace = F)
summary(fit.best)

```

Based on the AIC, the best model has predictors: `wt, am, qsec`
Now we can fit a parsimonious model. This model now gives a Adjusted Rsq of 83.3 with a low
p-value indicatiing the model significance.


5. Hypothesis test for model selection
This test also suggest that the best.fit model is significant than the full model
```{r}
anova(fit.cyl, fit.best, fit.full)
```

### Model Diagnoistics

* The model diagnostics shows that the best fit model has linear pattern, indicating the linearity 
was captured by the mode
* residuals appear normaly distributed as shows on the Normal-Q-Q plot as well
* The Scale-Location shows that the best fit model has approximately constant variance  
* Residual vs Leverage captures influetial points, in particular, there is no outliers in our model

<a href = 'https://data.library.virginia.edu/diagnostic-plots/'> You can read more here </a>

```{r}
par(mfrow = c(2,2))
plot(fit.best)
```

Sum of residual is also zero, all assumptions for a linear regression are adequately satisfied
```{r}
sum(fit.best$residuals)
```


## Conclusion
* Is automatic or manual transmission better for MPG?
This question can be answered based on the model, that Automatic transmission 
appears to consume fewer MPG with Manual transmission expected consume additional 2.9 MPG

* Quantify the MPG difference between manual and automatic transmission could be hard to answer as different models could predict different results


